---
title: "Hollywood Age Gap"
subtitle: "Cassie Minicucci, Katie Riley, Basil-Anne Stackpole"
format: 
  revealjs
highlight-style: a11y
execute:
  code-link: true
  echo: true
editor_options: 
  chunk_output_type: console
---
```{r}
#| echo: false

library(quarto)
library(readr)
library(tidytuesdayR)
library(dplyr)
library(ggplot2)
library(mdsr)
age_gaps <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv", show_col_types = FALSE)

```


## Introduction

* Dataset: Hollywood Age Gap
  + By Lynn Fisher, a web designer
  + Compiles the age differences between love interests in films
  
* 12 variables and 1177 observations before wrangling

* Questions: 
  + Are men typically the older person in the age gap, and has this changed over time?
  + Are certain directors more likely to have larger age gaps in their films than others?


## Relevant variables
* actor_1_age and actor_2_age: ages of both actors in a screened relationship
* character_1_gender and character_2_gender: genders of both characters in a screened relationship
* release_year: the year a film was released
* age_difference: the age gap between both partners (in years)
* director: the director of the movie


## Question 1 {.smaller}

Are men typically the older person in the age gap, and has this changed over time?

::: {.panel-tabset}

### Code

```{r}
#| label: fig3
#| fig-show: hide

custom_colors <- c("man" = "cyan3", "woman" = "deeppink3", "Other" = "gray")

older_actors <- ggplot(age_gaps, aes(release_year, color = character_1_gender))+
  geom_freqpoly() +
  scale_color_manual(values = custom_colors) +
  theme_bw()
older_actors
```

### Plot

```{r}
#| ref.label: fig3
#| echo: false 
```

:::

## Another visualization

::: {.panel-tabset}

### Code

```{r}
#| label: fig1
#| fig-show: hide

custom_colors <- c("man" = "cyan3", "woman" = "deeppink3", "Other" = "gray")

age_diff_plot <- ggplot(age_gaps, aes(release_year)) +
  geom_smooth(aes(y = actor_1_age, color = character_1_gender, linetype = "Actor 1")) +
  geom_smooth(aes(y = actor_2_age, color = character_2_gender, linetype = "Actor 2")) +
   scale_color_manual(values = custom_colors) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs( title = "Actor age gaps over time" ,
        x = "movie release year",
        y = "actor age") +
  theme_bw()
age_diff_plot
```

### Plot

```{r}
#| ref.label: fig1
#| echo: false 
```

:::

## Question 2
 
* Are certain directors more likely to have larger age gaps in their films than others?

### Data wrangling

```{r}
movies_1 <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv", show_col_types = FALSE)
movies_2 <- movies_1 %>% count(director)%>%
  arrange(desc(average_age_gap))
```

```{r}
movies_1 <- movies_1 %>%
  group_by(director) %>%
  mutate(average_age_gap = mean(`age_difference`)) 

```

```{r}
movies_usable <- subset(movies_1, director %in% c("Woody Allen", "John Glen", "Martin Scorsese", "Mike Newell", "Steven Spielberg", "Christopher Nolan", "David Fincher", "Dennis Dugan", "Guy Hamilton", "Mike Nichols"))

```

## Visualization

::: {.panel-tabset}

### Code

```{r}
#| label: fig2
#| fig-show: hide

director_plot <- ggplot(movies_usable, aes(x = director, y = age_difference, fill= director)) +
  geom_boxplot() +
  labs(title = "Typical Age Gaps in Hollywood Movies by Common Directors",
       x = "Director",
       y = "Age Difference")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = "none") #no legend bc it's obvious from axis labels

director_plot
```

### Plot

```{r}
#| ref.label: fig2
#| echo: false 
```

:::

## Summary statistics

::: {.panel-tabset}

### Summary stats

```{r}
movie_stats <- summary(movies_usable$age_difference)
print(movie_stats)

```

### Mean by director

```{r}
movies_usable_mean <- movies_usable %>%
  group_by(director) %>%
  summarize(average_age_gap = mean(`age_difference`), n = n()) %>%
  arrange(desc(average_age_gap))

print(movies_usable_mean)

```
>>>>>>> e3f1cf94cb75995b5d2092fc83f8614fe155ef0f

:::

## Conclusion

* Although we can't come to any causal conclusions due to this being an observational experiment, our data suggests that men are, in fact, typically the older person in an age gap.

* Additionally, directors like Woody Allen and John Glen seem to feature larger age gaps in their movies, whereas directors like Christopher Nolan and Mike Newell seem more likely to have smaller age gaps.
